Modeling the Linux page cache for accurate simulation of data-intensive applications

### 2020-12-15 paper_io_sim.tex

        To parametrize the simulators, we benchmarked the
        memory, local disk, remote disk (NFS), and network bandwidths
        (Table~\ref{table:benchmark}). Since \simgrid, and thus \wrench, currently only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.
            \begin{table}
            \centering
            \begin{tabularx}{\columnwidth}{ll
            >{\centering\arraybackslash}X
            >{\centering\arraybackslash}X
            >{\centering\arraybackslash}X}
            \toprule
                \multicolumn{2}{c}{Bandwidths}  & Cluster (real) & Python prototype & \wrench simulator\\
            \midrule
            \multirow{2}{*}{Memory}      & read  & 6860 & 4812 & 4812\\
                                         & write & 2764 & 4812 & 4812\\
            \multirow{2}{*}{Local disk}  & read  & 510  & 465  & 465\\
                                         & write & 420  & 465  & 465\\
            \multirow{2}{*}{Remote disk} & read  & 515  & -    & 445\\
                                         & write & 375  & -    & 445\\
            \multicolumn{2}{l}{Network}  & 3000  & -    & 3000\\
            \bottomrule
            \end{tabularx}
            \caption{Bandwidth benchmarks (MBps) and simulator configurations.
            The bandwidths used in the simulations are the average of the read and write bandwidths
            measured in the cluster.
            Remote disk and network accesses are not simulated in the Python simulator.}
            \label{table:benchmark}
            \end{table}

@@

    \section{Conclusion}
    \label{discussion}
    We designed a model of the Linux page cache and implemented it in the
    \simgrid-based \wrench library to simulate distributed applications.
    Evaluation results show that our model improves simulation accuracy
    substantially, reducing absolute relative simulation errors up to
    9$\times$ (see results of the single-threaded experiment). The
    availability of asymmetrical disk bandwidths in the forthcoming
    \simgrid release will further improve these results.
    Our page cache model is publicly available in the \wrench GitHub
    repository.

@@

    I/Os remain much more complex than is currently available in
    simulation models, which could substantially limit the accuracy of simulations.
    Relevant extensions to this work could include more
    accurate descriptions of anonymous memory usage in applications, 
    which strongly affects I/O times through writeback cache. File access patterns, 
    including random I/O, might also be worth including in application models,
    as they directly affect page cache content.
    % readahead and persistent storage could also be adeed
    % Simulation results could be made even more accurate by a deeper
    % investigation of page cache flushing and eviction order.

>>>>

        To parameterize the simulators, we benchmarked the
        memory, local disk, remote disk (NFS), and network bandwidths
        (Table~\ref{table:benchmark}). Since \simgrid, and thus \wrench, currently only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.
            \begin{table}
            \centering
            \begin{tabularx}{\columnwidth}{ll
            >{\centering\arraybackslash}X
            >{\centering\arraybackslash}X
            >{\centering\arraybackslash}X}
            \toprule
                \multicolumn{2}{c}{Bandwidths}  & Cluster (real) & Python prototype & \wrench simulator\\
            \midrule
            \multirow{2}{*}{Memory}      & read  & 6860 & 4812 & 4812\\
                                         & write & 2764 & 4812 & 4812\\
            \multirow{2}{*}{Local disk}  & read  & 510  & 465  & 465\\
                                         & write & 420  & 465  & 465\\
            \multirow{2}{*}{Remote disk} & read  & 515  & -    & 445\\
                                         & write & 375  & -    & 445\\
            \multicolumn{2}{l}{Network}  & 3000  & -    & 3000\\
            \bottomrule
            \end{tabularx}
            \caption{Bandwidth benchmarks (MBps) and simulator configurations.
            The bandwidths used in the simulations are the average of the read and write bandwidths
            measured in the cluster.
            Remote disk and network accesses are not simulated in the Python simulator.}
            \label{table:benchmark}
            \end{table}

@@

    \section{Conclusion}
    \label{discussion}
    We designed a model of the Linux page cache and implemented it in the
    \simgrid-based \wrench simulation framework to simulate the execution
    of distributed applications.
    Evaluation results show that our model improves simulation accuracy
    substantially, reducing absolute relative simulation errors by up to
    9$\times$ (see results of the single-threaded experiment). The
    availability of asymmetrical disk bandwidths in the forthcoming
    \simgrid release will further improve these results.
    Our page cache model is publicly available in the \wrench GitHub
    repository.

@@

    Not all I/O behaviors are captured by currently available simulation models,
    including the one developed in this work, 
    which could substantially limit the accuracy of simulations.
    Relevant extensions to this work include more
    accurate descriptions of anonymous memory usage in applications, 
    which strongly affects I/O times through writeback cache. File access patterns, 
    including random I/O, might also be worth including in the simulation models,
    as they directly affect page cache content.
    % readahead and persistent storage could also be adeed
    % Simulation results could be made even more accurate by a deeper
    % investigation of page cache flushing and eviction order.
